Finding shortest and closest vectors in a lattice of Voronoi's first kind

### 2014-04-21 paper.tex
minor

\section{Voronoi cells and relevant vectors}\label{sec:voron-cells-relev}
\newcommand{\calR}{\mathcal{R}}
The (closed) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ in $\reals^m$ is the subset of $\reals^m$ containing all points nearer or of equal distance (here with respect to the Euclidean norm) to the lattice point at the origin than to any other lattice point. The Voronoi cell is an $m$-dimensional convex polytope that is symmetric about the origin. %Here we will always assume the Euclidean norm (or 2-norm), so $\vor(\Lambda)$ contains those points nearer in Euclidean distance to the origin. %If $x \in \Lambda$ it follows that $\vor(\Lambda) + x$ is the subset of $\reals^n$ that is nearer to $x$ than any other lattice point in $\Lambda$. Figure \ref{lattices:fig:vorregion} is an example of the Voronoi cell.

>>>>

\section{Voronoi cells and relevant vectors}\label{sec:voron-cells-relev}
\newcommand{\calR}{\mathcal{R}}
The (closed) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ in $\reals^m$ is the subset of $\reals^m$ containing all points closer or of equal distance (here with respect to the Euclidean norm) to the lattice point at the origin than to any other lattice point. The Voronoi cell is an $m$-dimensional convex polytope that is symmetric about the origin. %Here we will always assume the Euclidean norm (or 2-norm), so $\vor(\Lambda)$ contains those points nearer in Euclidean distance to the origin. %If $x \in \Lambda$ it follows that $\vor(\Lambda) + x$ is the subset of $\reals^n$ that is nearer to $x$ than any other lattice point in $\Lambda$. Figure \ref{lattices:fig:vorregion} is an example of the Voronoi cell.

### 2014-04-21 talk/presentation.tex
added lattice figures
minor

\frame{
\begin{figure}[tp]
	\centering 
		\includegraphics[width=\linewidth]{figs/preslatticefigures-1.mps}
		\caption{A 2-dimensional lattice.}
		\label{fig:latfig1}
\end{figure}
}

@@

\frame{
\begin{figure}[tp]
	\centering 
		\includegraphics[width=\linewidth]{figs/preslatticefigures-1.mps}
		\caption{A 2-dimensional lattice.}
		\label{fig:latfig2}
\end{figure}
}

\frame{
\begin{figure}[tp]
	\centering 
		\includegraphics[width=\linewidth]{figs/preslatticefigures-3.mps}
		\caption{A 2-dimensional lattice.  There are 4 short vectors.}
		\label{fig:latfigandpacking1}
\end{figure}
}

\frame{
\frametitle{The Voronoi cell}
}

\frame{
PICTURE OF VORONOI CELL
}

\frame{
\frametitle{Relevant vectors}
}

\frame{
PICTURE OF RELEVANT VECTORS
}

\frame{
\frametitle{}
}

@@

\frame{
\frametitle{The closest lattice point problem}
ADD THIS
}

>>>>

\frame{
\begin{figure}[tp]
	\centering 
		\includegraphics{figs/lattice-1.mps}
		\caption{A 2-dimensional lattice.}
		\label{fig:latfig1}
\end{figure}
}

@@

\frame{
\begin{figure}[tp]
	\centering 
		\includegraphics{figs/lattice-1.mps}
		\caption{A 2-dimensional lattice.}
		\label{fig:latfig2}
\end{figure}
}

\frame{
\begin{figure}[tp]
	\centering 
		\includegraphics{figs/latticewithspherepacking-1.mps}
		\caption{A 2-dimensional lattice.  There are 4 short vectors.}
		\label{fig:latfigandpacking1}
\end{figure}
}

@@

\frame{ 
\frametitle{The Voronoi cell}

The \term{Voronoi cell} of a lattice $\Lambda \subset \reals^m$ is the subset of $\reals^m$ with points at least as close to the origin than to any lattice point,
\[
\vor(\Lambda) = \{ x \in \reals^m \mid \|x\| \leq \|x - y\|, y \in \Lambda \}.
\]

\frame{ 
\begin{figure}[tp]
	\centering 
		\includegraphics{figs/voronoicell-1.mps}
		\caption{A 2-dimensional lattice and its Voronoi cell.}
		\label{fig:latfigandpacking1} 
\end{figure}
}

\frame{ 
\frametitle{Relevant vectors}

The \term{relevant vectors} of a lattice $\Lambda$ are those which contribute a face to the Voronoi cell.

\frame{
\begin{figure}[tp] 
	\centering 
		\includegraphics{figs/relevantvectors-1.mps}  
		\caption{A 2-dimensional lattice with 6 relevant vectors.}
		\label{fig:latfigandpacking1}
\end{figure}
}

\frame{ 
\frametitle{Relevant vectors}
\begin{itemize}
\item The set of relevant vectors $\relevant(\Lambda)$ contains those $v \in \Lambda$ such that  
\[
\dotprod{v}{x} \leq \dotprod{x}{x}
\]
for all $x \in \Lambda$ with $x \neq v$ and $x \neq \zerobf$.  
\item The Voronoi cell can be defined using the relevant vectors, 
\[
\vor(\Lambda) = \{ x \in\reals^m \mid \|x\| \leq \|x - v\|, v \in\relevant(\Lambda) \}
\]
\item Short vectors are relevant vectors.
\end{itemize}
}

\frame{
\begin{figure}[tp] 
	\centering 
		\includegraphics{figs/relevantvectorsandspherepacking-1.mps}   
		\caption{A lattice with 6 relevant vectors and 4 short vectors.}
		\label{fig:latfigandpacking1}
\end{figure}
}

\frame{
\frametitle{Closest lattice points}

Given a lattice $\Lambda \subset \reals^m$ and a vector $y \in \reals^m$, a problem of interest is to find $x \in \Lambda$ such that
\[
\| y - x \|^2
\] 
is minimised.

\begin{itemize}
\item This is called the \term{closest lattice point problem} and a solution is called a \term{closest lattice point} to $y$
\end{itemize}
